### 2016-09-29 paper2/paper.tex
paper 2 template

\documentclass[letterpaper, twocolumn,11pt]{article}
\usepackage{epsfig,endnotes,amsmath,usenix}

\title{\textbf{Regressions And Stuff}}

\author{
\rm{Vahid Fazel-Rezai}\\
\rm{\texttt{vfazel@mit.edu}}
\and 
\rm{Akshay Ravikumar}\\
\rm{\texttt{akshayr@mit.edu}}
}

\date{May 6, 2016}

\begin{document}
% \begin{titlepage}
% \centering
% Matei Zaharia\\
% TR 10\\
% May 6, 2016
% \end{titlepage}

\maketitle
\section{Introduction}

In this paper, we discuss various methods of calculating regression. First, we discuss the benfits of using Stochastic Gradient Descent versus Batch Gradient Descent. Next, we analyze the performance of these methods in performing a Least Squares Linear regression. Finally, we discuss blah blah

\section{Overview}

\section{Gradient Descent}
\subsection{Batch Gradient Descent}
\subsection{Stochastic Gradient Descent}
\subsection{Effect of Finite Gradient}

\end{document}